Coursera Statistical Inference Project Part One
===============================================

Introduction
------------
The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. Set lambda = 0.2 for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

Objective
---------
Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.

The simulation
--------------
The next code runs a thousand simulations of 40 exponential(0.2)s and store the values in a matrix with 1000 columns and 40 rows. Each matrix element corresponds to a value of an exponential(0.2). The vector **sim.means** contains the means of the thousand simulations. We define **dat** as a data.frame of the vector **sim.means**. 
```{r}
## 1000 simulations of the mean of 40 exponentials(0.2)s 
sim.vectors <- replicate(1000, rexp(40,0.2), simplify="data.frame")
sim.means <- as.vector(colMeans(sim.vectors))
```

```{r, eval=TRUE, echo=FALSE, fig.width=4, fig.height=4}
library(ggplot2)
## Plot of the distribution of the means
dat <- as.data.frame(sim.means)
g <- ggplot(dat, aes(x = sim.means)) + 
 geom_histogram(alpha = .20, binwidth=.3, colour = "black", aes(y = ..density..))
g
```
1. Center of the distribution
-----------------------------
### Required
Show where the distribution is centered at and compare it to the theoretical center of the distribution.
### Answer
We know that the theoretical mean is equal to 1/lambda = 1/0.2 = 5. The value of the mean for the distribution of means of our 1000 simulations is equal to:
```{r}
sapply(dat, mean)
```
This value is very close to the theoretical value. So our distribution is centered around the theoretical mean as we expected.

2. Variance of the distribution
-------------------------------
### Required
Show how variable it is and compare it to the theoretical variance of the distribution.

### Answer
To evaluate the variance of the distribution we calculate the standard error of the distribution: 
```{r}
sapply(dat, sd)
```
And compare this value with the teoretical value for the normal distribution according with the central limit theorem (CLT) given by $latex \frac{\sigma}{\sqrt(40)}$, that in our case is equal to
```{r}
5/sqrt(40)
```
We can see that the theoretical and experimental values are very close.
3. Aproximation to a normal distribution
----------------------------------------
### Required
Show that the distribution is approximately normal.
### Answer
To show that the distribution is approximately normal we're going to represen the standard normal distribution over the normalized simulation data.
```{r, eval=TRUE, echo=FALSE, fig.width=4, fig.height=4}
## Plot of the distribution of the means
dat <- as.data.frame(sim.means)
# We normalize the simulation data and define the data frame
sim.means.norm <- (sqrt(40)/5)*(sim.means-5)
dat.norm <- as.data.frame(sim.means.norm)
# Plot of the normalized simulated data. Note that know they are centered around zero as expected
g <- ggplot(dat.norm, aes(x = sim.means.norm)) + 
 geom_histogram(alpha = .20, binwidth=.3, colour = "black", aes(y = ..density..))
# We superimpose the standard normal distribution over our data.
g <- g + stat_function(fun = dnorm, size = 1)
g
```

4. Coverage of the confidence interval for 1/lambda
---------------------------------------------------
### Required
Evaluate the coverage of the confidence interval for 1/lambda: $latex (\bar{X}) \pm 1.96 \frac{S}{\sqrt{n}}$.
### Answer
```{r}
sapply(dat, mean) + c(-1.96,1.96)*sapply(dat, sd)
```
